### 2016-09-21 HRI/Parts/System.tex

\subsection{Robotic Platform}
\subsection{Emotion Enrichment System}