### 2016-10-18 JournalInteractionStudies/Rebuttal2.txt
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.

General

Reviewer 1
*****************************************
One sentece to introduce next case study was added at the end of each case study

Reviewer 3
*****************************************
Case study titles were changed:
First Case study: Empirical Emotion Parameters
Second Case Study: Addition of Context
Third Case Study: Literature Emotional Parameters