### 2016-10-30 JournalInteractionStudies/Rebuttal2.txt
Changes in some parts of the article, main changes are described in Rebuttal2.txt

General
* IT was corrected Cañamero reference
* It was added the following article as reference: Relationship between color and emotion
* Introduction parapragh on Case study section was extended to include the conexion and create a better connection among the papers.

* One sentece to introduce next case study was added at the end of each case study

Reviewer 1
*****************************************

>>Q: First case study: There is no discussion of the results. Whilst the statistical analysis is now clearer, the authors should explain what has been learned in relation to theses results. Why were Happiness and Anger recognized? Do the authors have some hypothesis? Can any generalization be drawn?

>>Q: Second case study: What were the scenarios?
A: a link to a video, which shows the emotions, was added
>>Q: "However, if the conveyed emotion does not match the scene, the audience does not recognize it.": How can the design allow for such claim? Is this supported by the results? Is it possible that only the context was recognized regardless of what the robot was doing?
A: the sentence were changed, to "this opens the question about the influence that could have the human actor and the scene on the recognition rate"
>>Q: Third one: Why was the questionnaire changed half way?
It is well established that velocity is related to arousal. Why additional features are needed? The literature suggests that it is enough to discriminate between low and high velocity.
A: The re
Reviewer 3
*****************************************
Case study titles were changed:
First Case study: Empirical Emotion Parameters
Second Case Study: Addition of Context
Third Case Study: Literature Emotional Parameters