### 2016-10-30 JournalInteractionStudies/Journal-AISB.tex

\documentclass[man,floatsintext]{apa6}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{apacite}
\usepackage{rotating}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\usepackage{graphicx}
\usepackage{todonotes}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{url}
\usepackage{slashbox}
\newcommand{\co}{\cellcolor{gray!40}}

\title{Robots Showing Emotions: Emotion Representation with no bio-inspired Body}
\author{Julian M. Angel-Fernandez and Andrea Bonarini}
\affiliation{Department of Electronics, Information, and Bioengineering\\ Politecnico di Milano\\ Milan, Italy\\
email: \{julianmauricio.angel,andrea.bonarini\}@polimi.it}

%TODO (Done)change non-human like to non-antrhropomorphic
%TODO (Done)add a paragrahp at begining of section Case studies explaining the path followed
%TODO (Done) correct the paragraph written
%TODO (Done) make modification to case studies subtitles to give more information
%TODO (Done)review reference: . A. Russell and J.-A. Bachorowski et al., "Facial and vocal expressions of emotion," Annu. Rev. Psychol., vol. 54, no. 1, pp. 329-349, 2003.
%TODO Write a paragraph with this information
%TODO (Done)Add the relevant case study number to Table 2, 4, 6, 7, 11
%TODO Pg. 18, line 9. Add the word "respectively" at the end, since it could also be interpreted as "low and high arousal emotions are difficult to discriminate between each other".
%TODO (Done) correct Cañamero reference
%TODO (Done) find a reference for color 
%TODO (Done)make all the names for emotions expressed the same way (e.g. capital initial and italic, or only capital initial, or only italic, or nothing).

\abstract{Robots should be able to represent emotional states to interact with people as social agents. There are cases where robots cannot have bio-inspired bodies, for instance because the task to be performed requires a special shape, as in the case of home cleaners, package carriers, and many others. In these cases, emotional states have to be represented by exploiting movements of the body. In this paper, we present a set of case studies aimed at identifying specific values to convey emotion trough changes in linear and angular velocities, which might be applied on different non-anthropomorphic bodies. This work originates from some of the most considered emotion expression theories and from emotion coding for people. We show that people can recognize some emotional expressions better than others, and we propose some directions to express emotions exploiting only bio-neutral movement.
}

\shorttitle{Robots Showing Emotions}

\begin{document}
\maketitle
\section{Introduction}
\input{./Parts/introduction}
\section{Related Work}
\label{sec:related_work}
\input{./Parts/Emotion_Projection}
\section{System}
\label{sec:robotic_platform}
\input{./Parts/system}
\section{Case Studies}
\label{sec:cases}
\input{./Parts/case_studies}
\section{Conclusions and Further Work}
\input{./Parts/conclusions}
\bibliographystyle{apacite}
\bibliography{Bibliography,Biblography,BibloNew}
\end{document}