### 2017-03-24 RO-MAN-2017-2/Parts/description.tex

The description files allow the parametrization and extent the system to fit other applications is required. To standarized the mechanisms to describe these files, it was decided to create a language to describe them. This allows other researchers to create their own files and extended the language to fit their needs. It was decided to use EBNF to describe language.
\subsection{Emotion Description} 
The language created to describe emotion descriptors is the following:
\begin{small}

\begingrammar
%		
<emotion description>: "'\{'" "'emotion:'" <string>"','" "'observation:'" <string>"','" <action> ("','"<action>)* "'\}'".

<action>:<string>"': \{'" <description>"','" <actions affected> "'\}'".

<description>: "'description: \{'" <action name>"','";<emotion>"','" <parameter's type> "'\}'".

<action name>: "'emotionProfileAction:'" <string>.

<emotion>: "'emotionProfileEmotion:'" <string>.

<parameter's type>: "'movement\_parameter'".

<actions affected>: "'actions: \{'" <action parameter> ; ("','" <action parameter>)* "'\}'".

<action parameter>: <string>"':\{'" <reference>"','" <repetition>"','" <parameters> "'\}'".

<reference>: "'reference:'" <number>.

<repetition>: "'repetition:'" "'yes'" | "'no'".

<parameters>: "'parameters: ['" <parameter description> ("','" <parameter description>)* "']'".

<parameter description>: <movement parameter description>.

<movement parameter description>: "'\{time:'" <number> "', space:'" <number> "'\}'".

\endgrammar
\end{small}
%TODO give information regardin 1) why repetition 2) why a set of parameters 3) why use of time and space

\subsection{Action Message}
The language created to describe action Messages is the following:
\begin{small}
\begingrammar
%		
<action>: <simple action> <compound action> <context>.

<simple action>: "'\{'" <action header>"','" "'parameters: ['" <simple action parameters> "'] \}'".

<compound action>: "'\{'" <action header>"','" "'parameters:['" <compound action parameters> "']\}'".

<action header>: "'type:'" <action type>"','" "'name: '" <string>"','" <is primary>.

<simple action parameters>: "'\{'" <parameter header>"','" <parameter description> "'\}'".

<compound action parameters>:  <simple action parameters> ("','" <simple action parameters>)* .

<context>: "'\{'" <action type>"','" <emotion sync>"','" <action sync>"','" <is primary>"','" " 'information:'" <string>"','" " 'actions: '" <action> "'\}'".

<parameter header>: "'type:'" <parameter type>"','" " 'name:'" <string>.

<parameter description>: <parameter amplitude> |; <parameter circle> | <parameter landmark> | <parameter point> | <parameter speech> | <parameter square> | <parameter time>.

<parameter type>: "'mandatory\_parameter'" | ;"'optional\_parameter'".

<is primary>: "'yes'" | "'no'".

<emotion sync>: "'yes'" | "'no'".

<action type>: "'parallel\_node'" | "'sequential\_node'" | ;"'simple\_action'" | "'composite\_action'".

\endgrammar
\end{small}

%TODO give information regarding: primary and emotion synch