### 2017-03-26 RO-MAN-2017-2/Parts/description.tex

The description files allow the parametrization and extent the system to fit other applications is required. To standarize the mechanisms to describe these files, it was decided to create a language to describe them. This allows other researchers to create their own files and to extend the language to fit their needs. It was decided to use the Extended Backus-Naur Form (EBNF) to describe this language.
\subsection{Emotion Description} 
The language created to describe emotion descriptors is the following:
\begin{small}

\begingrammar
%		
<emotion description>: "'\{'" "'emotion:'" <string>"','" "'observation:'" <string>"','" <action> ("','"<action>)* "'\}'".

<action>:<string>"': \{'" <description>"','" <actions affected> "'\}'".

<description>: "'description: \{'" <action name>"','";<emotion>"','" <parameter's type> "'\}'".

<action name>: "'emotionProfileAction:'" <string>.

<emotion>: "'emotionProfileEmotion:'" <string>.

<parameter's type>: "'movement\_parameter'".

<actions affected>: "'actions: \{'" <action parameter> ; ("','" <action parameter>)* "'\}'".

<action parameter>: <string>"':\{'" <reference>"','" <repetition>"','" <parameters> "'\}'".

<reference>: "'reference:'" <number>.

<repetition>: "'repetition:'" "'yes'" | "'no'".

<parameters>: "'parameters: ['" <parameter description> ("','" <parameter description>)* "']'".

<parameter description>: <movement parameter description>.

<movement parameter description>: "'\{time:'" <number> "', space:'" <number> "'\}'".

\endgrammar
\end{small}
The attribute repetition is used to specify if the descriptors of an emotion parameters must be executed in an infinite loop. As it was already mentioned, the emotional parameters implemented for this version correspond to the move actions.

\subsection{Action Message}
The language created to describe action Messages is the following:
\begin{small}
\begingrammar
%		
<action>: <simple action> <compound action> <context>.

<simple action>: "'\{'" <action header>"','" "'parameters: ['" <simple action parameters> "'] \}'".

<compound action>: "'\{'" <action header>"','" "'parameters:['" <compound action parameters> "']\}'".

<action header>: "'type:'" <action type>"','" "'name: '" <string>"','" <is primary>.

<simple action parameters>: "'\{'" <parameter header>"','" <parameter description> "'\}'".

<compound action parameters>:  <simple action parameters> ("','" <simple action parameters>)* .

<context>: "'\{'" <action type>"','" <emotion sync>"','" <action sync>"','" <is primary>"','" " 'information:'" <string>"','" " 'actions: '" <action> "'\}'".

<parameter header>: "'type:'" <parameter type>"','" " 'name:'" <string>.

<parameter description>: <parameter amplitude> |; <parameter circle> | <parameter landmark> | <parameter point> | <parameter speech> | <parameter square> | <parameter time>.

<parameter type>: "'mandatory\_parameter'" | ;"'optional\_parameter'".

<is primary>: "'yes'" | "'no'".

<emotion sync>: "'yes'" | "'no'".

<action type>: "'parallel\_node'" | "'sequential\_node'" | ;"'simple\_action'" | "'composite\_action'".

\endgrammar
\end{small}